Spatial Analysis

### 2021-06-29 Scripts/spatial_analisis.Rmd
all species analysis

# Load interpolated data
# "Centropristis striata"
# 
historic_tif <- my_path("R","Partial/Interpolation","tif_Centropristis_striata.csv", read = T) %>% 
  mutate(state = str_to_sentence(state))

@@

Species <- c(
  "Scomber scombrus",
  "Doryteuthis (Amerigo) pealeii", #longfin squid
  "Peprilus triacanthus", # butterfish
  "Illex illecebrosus", #shortfin squid
  "Paralichthys dentatus", #summer flounder
  "Stenotomus chrysops", #Scoop"
  "Centropristis striata", # black sea bass
  "Pomatomus saltatrix", # Bluefish
  "Lopholatilus chamaeleonticeps", #golden tilefish
  "Caulolatilus microps", #blueline tilefish
  "Clupea harengus" # Atlantic herring (Jointly in State and Federal Waters)

  )

@@

state_lat <- historic_tif %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CO","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VI")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(order= letters[1:11])

@@

# print for notebook
head(historic_tif)
```

@@

 data_grid <- historic_tif %>% 
  # left_join(periods) %>% 
  # group_by(state,lat,lon,period) %>% 
  group_by(state,lat,lon,region) %>% 
  summarise(mean = mean(value,na.rm= T), .groups = "drop") %>% 
  rename(ID = state)

ggplot(data = subset(data_grid, !is.na(mean))) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = mean,
      col = mean
    )
  ) +
  geom_sf(data = land_sf, aes()) +
  facet_wrap(~ ID  +  region,
             ncol = 4) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_viridis("Mean Proportion", alpha = 0.8) +
  scale_color_viridis("Mean Proportion", alpha = 0.8) +
  my_ggtheme_p()
```

@@

*Notes:* Panel represents the **Spring.** survey. This computation considers the Overlapping of state waters.

@@

state_fit <- historic_tif %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"))

land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ label, nrow = 1) +
  geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                      size = 2,
                      box.padding = 0.10,
                      hjust = 1) +
  labs(x = "Latitude", 
       y = "Longitude") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "bottom") +
  theme(legend.key.width = unit(4,"line")) +
  ggsave(filename = "proportion_chg.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 12,
         height = 5
         )

@@

```{r area_map, eval = T, echo = T, fig.width = 10, fig.height=6}

state_fit <- historic_tif %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop")

@@

# group by state
state_fit <- historic_tif %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat)

# Plot me
ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "right") +
  theme(legend.key.width = unit(1,"line")) + 
  ggsave(filename = "area_plot_avg.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

>>>>

# Species data

# Load interpolated data
# "Centropristis striata"
# 
# historic_tif <- my_path("R","Partial/Interpolation","tif_Centropristis_striata.csv", read = T) %>% 
#   mutate(state = str_to_sentence(state))

# Multiple species

Species <- c(
  "Scomber scombrus",
  "Doryteuthis (Amerigo) pealeii", #longfin squid WE DONT HAVE IT
  "Peprilus triacanthus", # butterfish
  "Illex illecebrosus", #shortfin squid
  "Paralichthys dentatus", #summer flounder
  "Stenotomus chrysops", #Scoop"
  "Centropristis striata", # black sea bass
  "Pomatomus saltatrix", # Bluefish
  "Lopholatilus chamaeleonticeps", #golden tilefish
  "Caulolatilus microps", #blueline tilefish
  "Clupea harengus" # Atlantic herring (Jointly in State and Federal Waters)
  )

spp_paths <- list.files(my_path("R",extra_path = "Partial/Interpolation"))

# Some gibiris for names
taxon_list <- gsub("_"," ",spp_paths)
taxon_list <- gsub("tif ","",taxon_list)
taxon_list <- gsub(".csv","",taxon_list)

taxon_list <- taxon_list[taxon_list %in% Species]

# Now set a list of files to read
taxon_read <- paste0("tif_",taxon_list,".csv")
taxon_read <- gsub(" ","_",taxon_read)

taxon_read <- spp_paths[spp_paths %in% taxon_read]

taxon_read <- paste0(my_path("R",extra_path = "Partial/Interpolation"),taxon_read)

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_read, fread)) %>% 
  mutate(state = str_to_sentence(state))

@@

@@

state_lat <- historic_spp %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CO","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VI")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(order= letters[1:11])

@@

# print for notebook
head(historic_spp)
```

@@

 data_grid <- historic_spp %>% 
  group_by(state,lat,lon,region) %>% 
  summarise(sum = sum(value,na.rm= T), .groups = "drop") %>% # sum of all species
  group_by(state,lat,lon,region) %>% 
  summarise(mean = mean(sum,na.rm= T), .groups = "drop") # average of all years

ggplot(data = subset(data_grid, !is.na(mean))) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = mean,
      col = mean
    )
  ) +
  geom_sf(data = land_sf, aes()) +
  facet_wrap(~ state  +  region,
             ncol = 4) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_viridis("Mean Proportion", alpha = 0.8) +
  scale_color_viridis("Mean Proportion", alpha = 0.8) +
  my_ggtheme_p()
```

@@

*Notes:* Figure represents the **Spring** survey. This computation considers the Overlapping of state waters.

### Proportional change per species

@@

total_fited <- historic_spp %>% 
  group_by(year,region,spp) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,region,spp) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region,spp) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring")) %>% 
  mutate(spp = gsub(" ","\n",spp))

# The plot
land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_grid(spp ~ label) +
  labs(x = "Longitude", 
       y = "Latitude") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line")) +
  ggsave(filename = "proportion_chg_spp.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 14,
         height = 17
         )

@@

### Aggregated Proportional change

```{r area_map_aggregated, eval = T, echo = T, fig.width = 10, fig.height=6}

total_fited <- historic_spp %>% 
  group_by(year,region) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  filter(str_detect(region,"Spring"))

land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ label, nrow = 1) +
  labs(x = "Longitude", 
       y = "Latitude") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "bottom",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14) +
  theme(legend.key.width = unit(2,"line")) +
  ggsave(filename = "proportion_chg_agg.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 9,
         height = 5
         )

```{r area_map_aggregated, eval = T, echo = T, fig.width = 10, fig.height=6}

total_fited <- historic_spp %>% 
  group_by(year,region) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop")

@@

total_fited <- historic_spp %>% 
  group_by(year,region) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat)

# Plot me
ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "right") +
  theme(legend.key.width = unit(1,"line")) + 
  ggsave(filename = "area_plot_avg_agg.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )
